---
title: "Project2"
author: "Chuanni He"
date: "2022-10-03"
---

# Backgroun Information
This project is prepared by Aniket Anil Walimbe and Chuanni He. It created a vignette  for reading and summarizing data from one of the [Financial data API](https://polygon.io/docs/getting-started). 

# Requirements
To use the functions for interacting with the Financial API, We used the following packages:  
* `httr`: Tools for Working with URLs and HTTP.  
* `jsonlite`: A JSON parser/generator optimized for the web. Its main strength is that it implements a bidirectional mapping between JSON data and the most important R data types.  
* `ggplot2`: A system for declaratively creating graphics, based on The Grammar of Graphics.  

```{r message=FALSE, warning=FALSE}
# Required library and key information
key = "EByG3OEye377G2rI0LRah2CUjGplA7TB"
library(httr)
library(jsonlite)
library(ggplot2)
library(tidyverse)
```

# API Interaction Functions
This section talks about how we define the functions to interact with the Financial API, as well as some helper functions working with data manipulation.  
  
`Aggregates`  

Our first function is to obtain the stock price for a stock over a given date range in custom time window sizes. This function interacts with the `Aggregates (Bars)` endpoint of the API. `stock` parameter refers to the ticker of the referred company. For example, the stock ticker for Apple Inc. is `AAPL`. Either lower-case or capital letters are acceptable. The second parameter `multiplier` refers to the time multiplier for the acquired data. For example, if the `timespan` is `day`, then `multiplier` of `5` indicates that 5-minute bars will be obtained. `multiplier` should be integer values. The last parameter `timespan` refers to the size of the time window. The possible values are `minute`, `hour`, `day`, `week`, `month`, `quarter`, `year`. Either lower-case or capital letters are acceptable. This function will return the close price in the given time period between 2021-7-22 to 2022-7-22 with a maximum data length of 200 records. The returned data type is data frame.
```{r}
Aggregates = function(ticker,multiplier, timespan) {
  # generate the URL
  base = "https://api.polygon.io/v2/aggs/"
  Endpoint = "ticker"
  ticker = toupper(ticker)
  timespan = tolower(timespan)
  call = paste(base,Endpoint,"/",ticker,"/range/",multiplier,"/",timespan,
               "/2021-07-22/2022-07-22?adjusted=true&sort=asc&limit=200&apiKey=",
               key, sep = "")
  # Acquire the data
  get = GET(call)
  parsed = fromJSON(rawToChar(get$content))
  result = parsed$results
  colnames(result) = c("TradingVolumn", "VolumnWeightedAverage", 
                       "OpenPrice","ClosePrice", "HighestPrice",
                       "LowestPrice","Timestamp","Transactions")
  
  return(result)
}
```

`GetNews`  

The second function acquires the most recent news articles relating to a stock ticker. The summary of the article and the link to the original article are also included. Three parameters are included in this function. The `ticker` is the stock ticker of the company. Either lower-case or capital letters are acceptable. The `published` parameter refers to the date on witch the news is published greater than the specified value. The data type is string and the format should be `yyyy-mm-dd`. The third parameter y`limit` limits the number of news will be returned from the function. The data type is integer. This function returns a structured list including multiple information regarding each news such as publisher, author, title, keywords, etc.
```{r}
GetNews = function(ticker, published, limit) {
  ticker = toupper(ticker)
  # generate the URL
  base = "https://api.polygon.io/v2/reference/"
  Endpoint = "news"
  call = paste(base,Endpoint,"?ticker=",ticker,"&published_utc.gt=",published,
               "&limit=",limit, "&apiKey=", key, sep = "")
  # Acquire the data
  get = GET(call)
  parsed = fromJSON(rawToChar(get$content))
  # We removed the publisher information
  result = parsed$results[,-2]
  return(result)
}
```


# Data Exploration
Now that we can interact with a few of the endpoints of the Financial API, let’s get some data from them.  
First, let’s pull the stock price of Apple Inc. (the ticker is `AAPL`) and Amazon.com Inc. (the ticker is `AMZN`) during 2021-7-22 and 2022-7-22. We set the timeframe as 1 day (`multiplier=1`, `timespan="day"`), then, we call the function `Aggregates("aapl", 1, "day")` and `Aggregates("amzn", 1, "day")`. Then combine them together by adding a `Company` variable.
```{r}
Price_Apple = Aggregates("aapl", 1, "day")
Price_Amazon = Aggregates("amzn", 1, "day")

Price_Apple = data.frame(Company="Apple",Price_Apple)
Price_Amazon = data.frame(Company="Amazon",Price_Amazon)
Price = rbind(Price_Apple,Price_Amazon)
head(Price)
```
Plots
You should pull data from at least two calls to your obtaining data function (possibly combining them into one)
We have called the function twice to obtain the data for Apple and Amazon companies. We have then plotted the opening price for the stocks against the timestamp for doing various analysis like identifying the trends etc.

```{r}
ggplot(data = Price, aes(x = Timestamp, y = OpenPrice)) +
geom_line(aes(color = Company))

```
You should create at least one new variable that is a function of other variables :
Here, we have created a new variable "PriceVariation" which is the difference between the highest and lowest price of the stocks of a company per day. This variation is then plotted against the timestamp and following graph is obtaine. 

```{r}
priceUpdated <- Price %>%
  filter(Company == "Apple") %>%
  mutate(PriceVariation = (HighestPrice-LowestPrice)) %>%
  arrange(desc(PriceVariation))

ggplot(data = priceUpdated, aes(x = Timestamp, y = PriceVariation)) +
  geom_col()
```

```{r}
g <- ggplot(data = Price_Apple, aes(x = OpenPrice))
g + geom_histogram()
```



Next, let's create some contingency tables to summarize the key words for different types of companies in the news. We are interested in the key words in news for two types of companies: technology company and finance company. We selected `AAPL` and `AMZN` as technical company samples,  `JPM` and `WFC` as financial company samples. 
```{r}
News_AAPL = GetNews("aapl","2021-07-01",500)
wordvec_AAPL = tolower(unlist(News_AAPL$keywords))

News_AMZN = GetNews("amzn","2021-07-01",500)
wordvec_AMZN = tolower(unlist(News_AMZN$keywords))

News_JPM = GetNews("jpm","2021-07-01",500)
wordvec_JPM = tolower(unlist(News_JPM$keywords))

News_WFC = GetNews("wfc","2021-07-01",500)
worvec_WFC = tolower(unlist(News_WFC$keywords))

```
We grabbed the keywords in the first 500 news of the corresponding companies from 2021-07-01, save it as a string vector. Then we want to search for the keywords of `tech`, `investing`, and `markets` in the news from the four companies. We first create a `keywords` vector to save the word list. Next we search for the keywords in the corresponding string vectors, and create four contingency tables accordingly.

```{r}

keywords = c("tech", "investing","markets","eurozone")
wordcount_AAPL = wordvec_AAPL[wordvec_AAPL %in% keywords]
wordcount_AMZN = wordvec_AMZN[wordvec_AMZN %in% keywords]
wordcount_JPM = wordvec_JPM[wordvec_JPM %in% keywords]
wordcount_WFC = worvec_WFC[worvec_WFC %in% keywords]

table(wordcount_AAPL)
table(wordcount_AMZN)
table(wordcount_JPM)
table(wordcount_WFC)
```
It can be seen that there is a clear association between the keywords and company types. Two technical companies all includes keyword of `tech`, while either of the financial companies has fewer frequency of `tech`. On the other hand, financial terminology such as `eurozone` has higher frequencies in financial companies.
