---
title: "Project2"
author: "Chuanni He"
date: "2022-10-03"
---

# Backgroun Information
This project is prepared by Aniket Anil Walimbe and Chuanni He. It created a vignette  for reading and summarizing data from one of the [Financial data API](https://polygon.io/docs/getting-started). 

# Requirements
To use the functions for interacting with the Financial API, We used the following packages:  
* `httr`: Tools for Working with URLs and HTTP
* `jsonlite`: A JSON parser/generator optimized for the web. Its main strength is that it implements a bidirectional mapping between JSON data and the most important R data types.

```{r message=FALSE, warning=FALSE}
# Required library and key information
key = "EByG3OEye377G2rI0LRah2CUjGplA7TB"
library(httr)
library(jsonlite)
```

# API Interaction Functions
This section talks about how we define the functions to interact with the Financial API, as well as some helper functions working with data manipulation.  
  
`Aggregates`  

Our first function is to obtain the aggregate bars for a stock over a given date range in custom time window sizes.
```{r}
# timespan = (minute,hour,day,week,month,quarter,year)
Aggregates = function(stock,multiplier, timespan) {
  base = "https://api.polygon.io/v2/aggs/"
  Endpoint = "ticker"
  stock = toupper(stock)
  call = paste(base,Endpoint,"/",stock,"/range/",multiplier,"/",timespan,
               "/2021-07-22/2022-07-22?adjusted=true&sort=asc&limit=200&apiKey=",
               key, sep = "")
  
  get = GET(call)
  content = content(get)
  return(content)
}

stock = "aapl"
multiplier = 1
timespan = "day"
content = Aggregates(stock,multiplier,timespan)
```



```{r}
for (i in 1:length(content$results)) {
  print(content$results[[i]]$c)
}
```

